---
title: "autotest with R"
subtitle: ""
author: "Yalei Du"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: footer.css
    includes:
      after_body: footer.js
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

# autotest

- Repo: [https://github.com/casunlight/autotest](https://github.com/casunlight/autotest)

- Document: [http://autotest-docs.readthedocs.io/](http://autotest-docs.readthedocs.io/en/latest/R/index.html)

## Goal

1. Few test code 

1. Human reable error message 


---
# Installation

```bash
git clone https://github.com/casunlight/autotest.git
R CMD INSTALL autotest  # You may need to add the `sudo` prefix
```


---
# Guideline

- Baisc example

- Default error message 

- Automatic tracing

- Customize error message


---

# Baisc Example

```{r, eval=FALSE}
# submissions
x = 1
y = 2

# test code
library(autotest)
test_that('test x', {
  expect_equal(x, 1)
})
test_that('test y', {
  expect_equal(y, 1)
})
```

Mainly use the `test_that` and the `expect_equal` functions.

---

# Baisc Example

A `test_that` function call is a test block. It may contains multiple
expectations(represented as an `expect_equal` function call). 

Example:

```r
x = c(1, 2)
test_that('test x', {
  expect_equal(x[1], 1)
  expect_equal(x[2], 2)
})
```

However, you can simplified to a single expectation in this case: 

```r
test_that('test x', {
  expect_equal(x, c(1, 2))
})
```


---
class: center, inverse, middle

# Default error message 

### The default error message<sup>[1]</sup> is good enough

### **No need** extra settings

.footnote[
[1]: [Error Cases](http://autotest-docs.readthedocs.io/en/latest/R/4-example.html)
]

---
# Default error message 

- Variable Not Found


```r
library(autotest)
test_that('', {
  expect_equal(x, 1)
})
## AutoTestCaseError:
## Testing variable/expression:  x
## object 'x' not found
```

---

# Default error message 

- Variable Not Found


```r
library(autotest)
test_that('', {
  expect_equal(x, 1)
})
## AutoTestCaseError:
## Testing variable/expression:  x
## object 'x' not found
```

---

# Default error message 

- Function Not Found

```r
f <- 1
test_that('',{
  expect_equal(f(1), 1)
})
## AutoTestCaseError:
## Testing variable/expression:  f(1)
## could not find function "f"
```

---

# Default error message 

- Function Return Nothing

```r
f <- function(x) {}
test_that('',{
  expect_equal(f(1), 1)
})
## AutoTestCaseError:
## Testing variable/expression:  f(1)
## We expect your answer returns type "numeric", but it returns "NULL" instead.
## Do you forget to return something in your function definition?
```

---

# Default error message 

- Result Not Equal

```r
f <- function(x) ifelse(x > 5, x + 1, x)
test_that('',{
  for (i in 1:10){
    expect_equal(f(i), i + 1)
  }
})
## AutoTestCaseError:
## Testing variable/expression:  f(i)
## Your answer is 1, which is not equal to the correct answer 2
```

---

class: center, middle, inverse

# Read more cases here

## [Error Cases](http://autotest-docs.readthedocs.io/en/latest/R/4-example.html)


---

class: center, middle, inverse

# Automatic Tracing

---
# Automatic Tracing

```{r, eval=FALSE}
test_that('testing x', {
    expect_equal(x, 1)
})
## AutoTestCaseError:
{{## Testing variable/expression: x}}
## object ‘x’ not found
```

---
# Automatic Tracing

```{r, eval=FALSE}
test_that('testing f(i)', {
    expect_equal(f(i), i + 1)
})
## AutoTestCaseError:
{{## Testing variable/expression: f(i)}}
## could not find function "f"
```


---
# Automatic Tracing

When error occurs, the error message tells you what variable/expression
we are testing. This feature is trying to **LOCATE** the error.

- Almost all the `expect_XXX` functions support this feature.

- Always put the tested object on the first argument

.pull-left[
### Right
```r
expect_equal(x, 1)
```
]

.pull-right[
### Wrong
```r
expect_equal(1, x)
```
]


---
# Automatic Tracing: drawback


```{r, eval=FALSE}
f <- function(x) 1
test_that('testing f(i)', {
  i = 1
  expect_equal(f(i), i + 1)
})
## AutoTestCaseError:
## Testing variable/expression: f(i)
## Your answer is 1, which is not equal to the correct answer 2
## The maximum tolerance is 1e-05
```

> Testing variable/expression: f(i)

What is **i** here?

---

# Automatic Tracing: close it

```{r, eval=FALSE}
f <- function(x) 1
test_that('testing f(i)', {
  i = 1
{{ expect_equal(f(i), i + 1, trace=FALSE)}}
})
## AutoTestCaseError:
## Your answer is 1, which is not equal to the correct answer 2
## The maximum tolerance is 1e-05
```

No trace here. Where this error comes from?

---
# Add Self-defined Error Message

```{r, eval=FALSE}
f <- function(x) 1
test_that('testing f(i)', {
  i = 1
{{ registerPreMsg('In testing f(%s):', i)}}
  expect_equal(f(i), i + 1, trace=FALSE)
})
## AutoTestCaseError:
## In testing f(1):
## Your answer is 1, which is not equal to the correct answer 2
## The maximum tolerance is 1e-05
```

**registerPreMsg**: register error message before the default error

---
class: inverse, middle, center

# Customiz Error Message


---
# Customiz Error Message

- [registerPreMsg](): **register** error message **before** the default error


- [registerPostMsg](): **register** error message **after** the default error

- [addPreMsg](): **append** error message **before** the default error


- [addPostMsg](): **append** error message **after** the default error


---
# Exmaple:

```{r eval=FALSE}
x = 1
test_that('', {
{{ registerPreMsg('Here is the default error message:')}}
{{ registerPostMsg('----------------------')}}
  expect_equal(x, 2)
})
## AutoTestCaseError:
## Testing variable/expression:  x
{{## Here is the default error message:}}
## Your answer is 1, which is not equal to the correct answer 2
## The maximum tolerance is 1e-05
{{## ----------------------}}
```

---
# Working Space: Bad Case

The customized error message **works in the whole test block**(the `test_that` function call)


```{r eval=FALSE}
x = c(1, 3)
test_that('', {
{{ registerPreMsg('Test the first element of x:')}}
  expect_equal(x[1], 1)  # pass
  expect_equal(x[2], 2)
})
## AutoTestCaseError:
## Testing variable/expression:  x[2]
{{## Test the first element of x:}}
## Your answer is 3, which is not equal to the correct answer 2
## The maximum tolerance is 1e-05
```

Customized message show up when the second `expect_equal` call raise an error.

---
# Working Space: Good Case

Clear after checking the first element.

```{r eval=FALSE}
x = c(1, 3)
test_that('', {
  registerPreMsg('Test the first element of x:')
  expect_equal(x[1], 1) 
{{ registerPreMsg('')}}
  expect_equal(x[2], 2)
})
```

---
# Working Space

```{r eval=FALSE}
x = 1; y = 2
test_that('', {
{{ registerPreMsg('Test the first element of x:')}}
  expect_equal(x, 1)  # pass
})
test_that('', {
  expect_equal(y, 1)
})
## AutoTestCaseError:
## Testing variable/expression:  y
## Your answer is 1, which is not equal to the correct answer 2
## The maximum tolerance is 1e-05
```

Customized message in the first test block does not affect the other test blocks.

---
# Customized Error

Set `suppressErr=TRUE` to disable the default error, use `registerPostMsg` to add customized error message.

```{r eval=FALSE}
test_that('', {
    registerPostMsg('You answer is wrong')
    expect_equal(x, 2, suppressErr=TRUE)
})
## AutoTestCaseError:
## Testing variable/expression:  x
## You answer is wrong
```


---
# Customized Error

Answer:
```{r}
library(ggplot2)
p_scatterplot <- ggplot(data = cars, aes(x = speed, y = dist))
p_scatterplot <- p_scatterplot + geom_point()
```

Test Code:
```{r eval=FALSE}
library(autotest)
test_that("test p_scatterplot", {
  expect_is(p_scatterplot, 'ggplot')
  
  registerPostMsg('Did you map the `speed` column on the x axis?')
  expect_true(unlist(p_scatterplot)$labels.x == 'speed', 
              suppressErr=T)
  registerPostMsg('Did you map the `dist` column on the y axis?')
  expect_true(unlist(p_scatterplot)$labels.y == 'dist', 
              suppressErr=T)
  registerPostMsg('Did you add a point layer?')
  expect_is(layer_grob(p_scatterplot)[[1]], 'points', suppressErr=T)
})
```


---
class: inverse, middle, center

# Expect Functions


---
# Expect Functions

- [expect_equal](): compare two objects

- [expect_true](), [expect_false](), [expect_na](): logical

- [expect_is](), [expect_length](): type and length

- [expect_match](), [expect_output](), [expect_error](): strings

---
# expect_equal

Support `integer`, `character`, `logical`, `double`, `factor`, `list`, `matrix` and `data.frame`.

Here are the steps of comparing two objects:

- Test class

- Test length/dimension

- Test values

So [expect_equal(rep(1, 10), 1)]() will raise an error because of the length diff.

---
class: center, middle

# [Details of exepct_equal](http://autotest-docs.readthedocs.io/en/latest/R/2-expect_functions.html#numeric)


---
# exepct_equal

Finally, `exepct_equal` can not compare missing values, unless two objects are identical.

```{r eval=FALSE}
x = c(1, NA, 3)
names(x) = c('a', 'b', 'c')
expect_equal(x, c(1, NA, 3))
## AutoTestCaseError:
## Testing variable/expression:  c(1, NA, 3)
## Your answer conatins missing values NA, please check again.
```

```{r eval=FALSE}
names(x) = NULL
expect_equal(x, c(1, NA, 3))  # pass, because they are identical
```



---
# expect_XXX(logical)

[expect_true & expect_false & expect_na](): these functions are checking whehter the objects are TRUE/FALSE/NA.


**Note**: they support testing a vector of logical values.

```{r eval=FALSE}
x = 1:10
expect_true(x >= 1)
```

In the [testthat]() package, you have to write:
```r
testthat::expect_true(all(x >= 1))
```


---
# expect_is & expect_length

- [expect_is(x, 'integer')]() is equivalent to [expect_equal(class(x), 'integer')](). 

- [expect_length(x, 10)]() is equivalent to [expect_equal(length(x), 10)]().

---
# Example

If you want to test whether [y]() is a vector of TRUE with length 10:

```r
test_that('', {
    expect_length(y, 10)
    expect_true(y)
})

# alternatively
test_that('', {
    expect_equal(y, rep(TRUE, 10))
})
```

---
## expect_match

Whether string matches a pattern.

```{r, eval=FALSE}
expect_match('abc', 'a')  

expect_match('abc', '[a-z]{3}')  # using regular expression

expect_match('abc', 'A', ignore.case=T)  # case insensitive

# exactly match, not using regular expression
expect_match('a(b', '(b', fixed=T)  
```

---
# expect_output

Capture the output and check whether it matches a pattern.

```{r eval=FALSE}
expect_output(print('abc'), 'a')  

expect_output(cat('abc'), '[a-z]{3}')  # using regular expression

expect_output(cat('abc'), 'A', ignore.case=T) # case insensitive
```


---
# expect_error

Check whether the expression raise an error.


```{r eval=FALSE}
expect_error(1 - '1')

# check the error message matches our string
expect_error(1 - '1', 'non-numeric argument to binary operator')

# check the error message matches our pattern(regular expression)
expect_error(1 - '1', 'non-numeric.*')
```


---
# Summarize

1. Easy to learn

1. Works well without any extra settings

1. Able to customize

---
class: center, middle

# That's All!

Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).

The chakra comes from [remark.js](https://remarkjs.com), [**knitr**](http://yihui.name/knitr), and [R Markdown](https://rmarkdown.rstudio.com).
